---
title: 'MIS-64018: Assignment 3'
author: "Steve Spence"
date: "9/18/2019"
output: word_document
---

## Assignment 3, Problem 1:

First, we will load the "lpSolve" and "lpSolveAPI" to solve the problem outlined in parts A and B. This will load the appropriate tools needed to solve this problem.

```{r}

# Load the "lpSolveAPI" package into the R environment.

library(lpSolve)
library(lpSolveAPI)

```

First, we need to create a linear programming object. Based on our outline in steps A and B, this program will have 9 decision variables, 12 constraints, and minimum boundary conditions for each variable.

```{r}

# Begin the lp problem with 12 constraints and 9 decision variables.

lprec <- make.lp(12,9)

```

Next, we must set the objective function into our program. As a default, the program sets the objective function to find the minimum; however, in this case we will need to find the maximum profits, so we will also change that to look for the maximum value for the objective function.

```{r}

# Set the objective function for the problem.

set.objfn(lprec, c(420,420,420,360,360,360,300,300,300))

# Change the direction to set maximization

lp.control(lprec, sense = "max")

```

All of the constraint values will also need to be added into the program. These are the capacity, storage, sales, and capacity usage constraints previously defined in parts A and B. These will be input using the "set.row" command and defining which index to put the values at. This will prevent us from inputting a lot of "0" values manually.

```{r}

# Set the constraint values row by row

# Capacity constraints:

set.row(lprec, 1, c(1,1,1), indices = c(1,4,7))
set.row(lprec, 2, c(1,1,1), indices = c(2,5,8))
set.row(lprec, 3, c(1,1,1), indices = c(3,6,9))

# Storage constraints:

set.row(lprec, 4, c(20,15,12), indices = c(1,4,7))
set.row(lprec, 5, c(20,15,12), indices = c(2,5,8))
set.row(lprec, 6, c(20,15,12), indices = c(3,6,9))

# Sales constraints:

set.row(lprec, 7, c(1,1,1), indices = c(1,2,3))
set.row(lprec, 8, c(1,1,1), indices = c(4,5,6))
set.row(lprec, 9, c(1,1,1), indices = c(7,8,9))

# Capacity usage constaints:

set.row(lprec, 10, c(900,900,900,-750,-750,-750), indices = c(1,4,7,2,5,8))
set.row(lprec, 11, c(450,450,450,-900,-900,-900), indices = c(2,5,8,3,6,9))
set.row(lprec, 12, c(450,450,450,-750,-750,-750), indices = c(1,4,7,3,6,9))

```

Now, we will need to set the constraint values from the problem statement. In this case, these values correspond to capacity and storage constraints at each plant, sales constraints for each size, and proportion of capacity usage for each plant.

```{r}

# Set the right hand side values

rhs <- c(750,900,450,13000,12000,5000,900,1200,750,0,0,0)
set.rhs(lprec, rhs)

```

Next, we need to define the inequality values for the problem. Majority of these are going to be less than or equal to since the outlined problem stated the maximum capacity constraints.

```{r}

# Set the constraint type

set.constr.type(lprec, c("<=","<=","<=","<=","<=","<=","<=","<=","<=","=","=","="))

```

For this problem, all values must be greater than 0.

```{r}

# Set the boundary condiiton for the decision variables

set.bounds(lprec, lower = rep(0, 9))

```

This set of code will name the decision variables and the constraints for the model.

```{r}

# Set the names of the rows (constraints) and columns (decision variables)

lp.rownames <- c("Plant 1 Capacity", "Plant 2 Capacity", "Plant 3 Capacity", "Plant 1 Storage", "Plant 2 Storage", "Plant 3 Storage", "Large Sales", "Medium Sales", "Small Sales", "Plant 1 and 2 Usage", "Plant 2 and 3 Usage", "Plant 1 and 3 Usage")
lp.colnames <- c("Plant 1L", "Plant 2L", "Plant 3L", "Plant 1M", "Plant 2M", "Plant 3M", "Plant 1S", "Plant 2S", "Plant 3S")
dimnames(lprec) <- list(lp.rownames, lp.colnames)

```

This command should return the linear program outline before executing the code, so we can determine if all the values are correct.

```{r}

# Return the linear programming object to ensure the values are correct

lprec

# The model can also be saved to a file

write.lp(lprec, filename = "Assignment_2_Problem3C.lp", type = "lp")

```

The following code will now look for an optimal solution. If it returns a "0" value then that means the model has found an optimal solution.

```{r}

# Solve the linear program

solve(lprec)

```

The model returned a "0", so it has found an optimal solution to the problem.

The function below will return what the maximum value for the objective function will be.

```{r}

# Review the objective function value

get.objective(lprec)

```

In this case, the maximum profits that can be achieved with these constraints is $696,000 per day.

Next, we will return the values of the decision variables to decide how many units of each type of product the plants should make.

```{r}

# Get the optimum decision variable values

get.variables(lprec)

```

Optimum decision variable values from the model:

Plant 1, Large: 516.67 units/day
Plant 2, Large: 0 units/day
Plant 3, Large: 0 units/day
Plant 1, Medium: 177.78 units/day
Plant 2, Medium: 666.67 units/day
Plant 3, Medium: 0 units/day
Plant 1, Small: 0 units/day
Plant 2, Small: 166.67 units/day
Plant 3, Small: 416.67 units/day

The following two segments of code will tell us where our values fall within the constraints, as well as return the surplus between the constraint and the actual value from the constraints.

```{r}

# Get the constraint values for the problem

get.constraints(lprec)

```

```{r}

# Review the surplus for each constraint

get.constraints(lprec) - rhs

```

## Assignment 3, Problem 2:

The "get.sensitivity.rhs()" function will be used to identify the shadow prices for the right hand side of the formulated linear programming problem.

```{r}

get.sensitivity.rhs(lprec)

```

The "get.sensitivity.obj()" function will be used to identify the reduced cost values for the objective function of the formulated linear programming problem.

```{r}

get.sensitivity.obj(lprec) 

```

## Dual Problem:


```{r}

# Begin the dual lp problem with 9 constraints and 12 decision variables.

lprec2 <- make.lp(9,12)

```


```{r}

# Set the objective function for the problem.

set.objfn(lprec2, c(750,900,450,13000,12000,5000,900,1200,750,0,0,0))

# Change the direction to set minimization

lp.control(lprec2, sense = "min")

```


```{r}

# Set the constraint values row by row

set.row(lprec2, 1, c(1,20,1,900,450), indices = c(1,4,7,10,12))
set.row(lprec2, 2, c(1,15,1,900,450), indices = c(1,4,8,10,12))
set.row(lprec2, 3, c(1,12,1,900,450), indices = c(1,4,9,10,12))

set.row(lprec2, 4, c(1,20,1,-750,450), indices = c(2,5,7,10,11))
set.row(lprec2, 5, c(1,15,1,-750,450), indices = c(2,5,8,10,11))
set.row(lprec2, 6, c(1,12,1,-750,450), indices = c(2,5,9,10,11))

set.row(lprec2, 7, c(1,20,1,-900,-450), indices = c(3,6,7,11,12))
set.row(lprec2, 8, c(1,15,1,-900,-450), indices = c(3,6,8,11,12))
set.row(lprec2, 9, c(1,12,1,-900,-450), indices = c(3,6,9,11,12))

```


```{r}

# Set the right hand side values

rhs2 <- c(420,420,420,360,360,360,300,300,300)
set.rhs(lprec2, rhs2)

```


```{r}

# Set the constraint type

set.constr.type(lprec2, c(">=",">=",">=",">=",">=",">=",">=",">=",">="))

```


```{r}

# Set the boundary condiiton for the decision variables

set.bounds(lprec2, lower = c(0,0,0,0,0,0,0,0,0,-1.00e+30,-1.00e+30,-1.00e+30))

```


```{r}

# Solve the linear program

solve(lprec2)

```


```{r}

# Review the objective function value

get.objective(lprec2)

```


```{r}

# Get the optimum decision variable values

get.variables(lprec2)

```

```{r}

get.sensitivity.rhs(lprec2)

```

```{r}

get.sensitivity.obj(lprec2)

```

